\Huge \textbf{Hybrid rendering}\\ \huge using Radeon Rays

### 2018-11-04 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on paper (only one major piece left)

\section{Getting started}
\subsection{Building using CMake}

@@

The first main part that needs to be implemented to render is a window. As already outlined in the research part there are a lot of libraries out there. For this implementation SDL was chosen, simply for the reason that creating an SDL windowing context with OpenGL is a breeze. SDL expects the programmer to supply what version of OpenGL will be coupled to this window, what rendering mode will be used (e.g. immediate vs core) and the settings of the window. In our case we'll use OpenGL 3.3 seeing it was the first modern implementation of SDL and consecutively the easiest to comprehend, we'll also be using core mode seeing we need vertex buffer abilities.

\subsection{OpenGL rendering context}
Once the window is in place the programmer is responsible for querying the accompanying OpenGL context.

\subsection{OpenGL function bindings}
Once the window and accompanying context is in place the programmer can move on to the OpenGL loading. For this implementation Glad was used. To get glad to work in this solution the respective package has to be created from their website \href{https://glad.dav1d.de/}. For our purpose I selected the GL 3.3 package in core functionality mode. After the package selection has been done you should receive the source files. These can then be compiled into a solution by making use of CMake, to then consecutively build those into lib files. The final step was to link both the includes and libs into my solution. A simple call to gladLoadGLLoader(SDL\_GL\_GetProcAddress) later and it's done.\newline\par

\pagebreak
\subsection{Shader loading}

One of the major things that make the graphics pipeline modular is programmable shaders. Shaders are mostly written in a pseudo c-language like GLSL/HLSL. This language is relative to the underlying rendering api that is being used. The rendering api is therefore in charge for the shader compilation. The only obstacle that the host program has to overcome is loading the source from the shader file. In this implementation the default cpp iostream was used, but this can be tailored to the needs of your engine/implementation. Once the shader source has been loaded it needs to be compiled and hooked to an OpenGL object. In this implementation a clear distinction between a shader and a shader program was made, where a shader is seen as the smallest building block and a shader program as a collection of those blocks.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{shader.jpg}
	\caption{Snippet from shader loader}
\end{figure}

@@

\begin{itemize}
	\item{ProcessNode}\newline\par
	ProcessNode as the name suggest goes over the current node and recursively calls the ProcessMesh on its meshes. Once all the meshes within this node have been handled it goes and calls ProcessNode on its children.
	\item{ProcessMesh}\newline\par
	ProcessMesh has three major tasks the first one being fetching all the vertices of the mesh things like: storing position, normal data and texture coordinates. Secondly it makes sure all the indices are being loaded in before finally going over to the materials where it calls the LoadMaterialTextures function on each of its materials.
	\item{LoadMaterialTextures}\newline\par
	Finally LoadMaterialTextures then takes in the material and the type of texture that we are trying to fetch, it then checks the 3D object format for that type of format and stores it respectively. As an added optimization it keeps track of double textures, so it won't load the same texture twice.
\end{itemize}

@@

\subsection{Coordinate systems}
Although you can render without a proper coordinate system in place, the implementation needed world space coordinates rather than screen space. \newline\par
So for this implementation a transform component was set in place. This transform component stands in relationship with the GameObject. This relationship allows us to add childs to the GameObject while maintaining the correct local transforms in the component. If the position of a said object was changed we go ahead and rebuild our worldmatrix while keeping in check if the object had a parent.\newline\par
The next part is the camera component, the camera component is in charge of holding the viewprojectionmatrix. Meaning that it is also in charge of holding the camera position and its settings.\newline\par
The final part to the puzzle is the mesh draw component. To create a meshdrawcomponent the programmer has to provide it with a mesh filter which is loaded by the model loading utilities mentioned prior. Then the final assembly is to build the worldviewprojectionmatrix (by multiplying the world, view and projection matrices with each other) and send it to the shader.

\subsection{Deferred shading}

>>>>

@@

The first main part that needs to be implemented to render is a window. As already outlined in the research part there are a lot of libraries out there. For this implementation SDL was chosen, simply for the reason that creating an SDL windowing context with OpenGL is a breeze. SDL expects the programmer to supply what version of OpenGL will be coupled to this window, what rendering mode will be used (e.g. immediate vs core) and the settings of the window. In our case we'll use OpenGL 3.3 seeing it was the first modern implementation of SDL and consecutively the easiest to comprehend, we'll also be using core mode seeing we need vertex buffer abilities. Once the window is in place the programmer is responsible for querying the accompanying OpenGL context.

\subsection{OpenGL function bindings}
Once the window and accompanying context is in place the programmer can move on to the OpenGL loading. For this implementation Glad was used. To get glad to work in this solution the respective package has to be created from their website \url{https://glad.dav1d.de/} For our purpose I selected the GL 3.3 package in core functionality mode. After the package selection has been done you should receive the source files. These can then be compiled into a solution by making use of CMake, to then consecutively build those into lib files. The final step was to link both the includes and libs into my solution. A simple call to gladLoadGLLoader(SDL\_GL\_GetProcAddress) later and it's done.\newline\par

\subsection{Shader loading}
One of the major things that make the graphics pipeline modular is programmable shaders. Shaders are mostly written in a pseudo c-language like GLSL/HLSL. This language is relative to the underlying rendering api that is being used. The rendering api is therefore in charge for the shader compilation. The only obstacle that the host program has to overcome is loading the source from the shader file. In this implementation the default cpp iostream was used, but this can be tailored to the needs of your engine/implementation. Once the shader source has been loaded it needs to be compiled and hooked to an OpenGL object. In this implementation a clear distinction between a shader and a shader program was made, where a shader is seen as the smallest building block and a shader program as a collection of those blocks.

\subsection{Hello Triangle}
With OpenGL and SDL established it is time to test the renderer by doing to famous hello triangle test. The first step for the programmer is to supply the vertices of the triangle, seeing there are still no coordinate systems in place it is ok to define those vertices in normalized device coordinates.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{ndc.png}
	\caption{Normalized device coordinates for our triangle}
\end{figure}

Once the vertex data is defined the programmer needs to send it of to the first stage of the rendering pipeline (vertex shader). This means that the programmer should first get those vertices on the GPU, in modern OpenGL this is done by requesting a piece of memory called a vertex buffer object. After binding this buffer and providing it from its buffer data the object should be set in place.\newline\par

Modern OpenGL requires the graphics programmer to atleast supply his own vertex and fragment shader, this is in contrary to DirectX where a default one is used if none supplied. Our vertex shader will consist of the vertex layout and the passing of our vertices position to the OpenGL system value gl\_Position. Under vertex layout the programmer provides how his vertices store the data needed in the shader, this also needs to be done in the host program through the use of glVertexAttributePointer. The aforementioned gl\_Position is there to provide a means of communication to specify where the vertex should appear on screen.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{vertexattributepointer.png}
\end{figure}

Then the final piece of the puzzle is to supply our pipeline of a fragment shader. The fragment shader in itself is quite simple only requiring us to specify the final fragments color, this can of course be any arbitrary value. After shader compilation, binding our shader, binding our vertex buffer and a final call to drawtriangles later a triangle appeared on screen.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{hellotriangle.png}
\end{figure}

\pagebreak
\subsection{Coordinate systems}
Although you can render without a proper coordinate system in place, the implementation needed world space coordinates rather than screen space. \newline\par
So for this implementation a transform component was set in place. This transform component stands in relationship with the GameObject. This relationship allows us to add childs to the GameObject while maintaining the correct local transforms in the component. If the position of a said object was changed we go ahead and rebuild our worldmatrix while keeping in check if the object had a parent.\newline\par
The next part is the camera component, the camera component is in charge of holding the viewprojectionmatrix. Meaning that it is also in charge of holding the camera position and its settings.\newline\par
The final part to the puzzle is the mesh draw component. To create a meshdrawcomponent the programmer has to provide it with a mesh filter which is loaded by the model loading utilities mentioned prior. Then the final assembly is to build the worldviewprojectionmatrix (by multiplying the world, view and projection matrices with each other) and send it to the shader.

@@

\begin{itemize}
	\item{ProcessNode} as the name suggest goes over the current node and recursively calls the ProcessMesh on its meshes. Once all the meshes within this node have been handled it goes and calls ProcessNode on its children.
	\item{ProcessMesh} has three major tasks the first one being fetching all the vertices of the mesh things like: storing position, normal data and texture coordinates. Secondly it makes sure all the indices are being loaded in before finally going over to the materials where it calls the LoadMaterialTextures function on each of its materials.
	\item{LoadMaterialTextures} then takes in the material and the type of texture that we are trying to fetch, it then checks the 3D object format for that type of format and stores it respectively. As an added optimization it keeps track of double textures, so it won't load the same texture twice.
\end{itemize}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{modelandtexture.png}
\end{figure}

@@

\subsection{Deferred shading}
With a standard rendering pipeline in place it was time to go ahead and turn that pipeline into a deferred one. The steps beforehand did also not cover lighting effects for the simple reason that they would need to be re-written once going deferred.\newline\par
The major implementation difference now is that the programmer needs to define his/her off screen render buffers (e.g. position buffer, normal buffer, etc.). First the programmer has to define what buffers are stored in the gbuffer object (frame buffer object). With the gbuffer object defined it's time to start moving on to the fill shader.\newline\par
The fill shader is tasked with filling the gbuffers with all the geometry information respectively, so if the implementation uses four buffers the fill shader will be tasked with filling those four buffers. The fill shader is a shader program consisting of two shaders first a simple vertex shader that passes the information to the next stages and then a fragment shader who does all the heavy lifting and fills all the buffers. Then all the geometry gets rendered using this shader. After the bulk of the rendering is done we unbind the gbuffers (that are now filled with geometry information) and we move on to the lightpass.\newline\par

The light pass shader is in charge of lighting and rendering the final full screen quad. It speaks for itself that before going into this lightpass we need to make sure the g-buffer is no longer bound as the active rendertarget and that the g-buffer textures can now be sent off to the lightpass to do its lighting calculations. Here again the implementation makes use of two shaders a very simple vertex shader that effectively passes texture coordinates and vertex position to the fragment stage. The fragment shader in this case will now resemble a traditional shader apart from the fact that it needs to sample all of its information from the geometry render buffers (g-buffer).\newline\par

The final piece of code is in charge of rendering a full screen quad with the combined g-buffer information (through the use of the light pass) as its texture.
